---
title: "Discovering tables in Manifold via ODBC"
author: "Michael Sumner"
date: "20 January 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 4, fig.height = 5)
library(manifoldr)
library(RODBC)
library(raster)
```


In this Project "Components.Map" I have a number of components, to try
different kinds of queries. I am using R with the CRAN package `RODBC`
and a custom function  `odbcConnectManifold`. (See Notes). 

```{r}
library(RODBC)  ## install.packages("RODBC")
library(manifoldr) ## https://github.com/mdsumner/manifoldr
# or just source the function in the Notes below
mapfile <- "C:/temp/Components.map"
con <- odbcConnectManifold(mapfile)
print(con)  ## summary of the connection
```

```{r, echo = FALSE, eval = TRUE}
library(manifoldr)
library(RODBC)
library(raster)
colors <- sqlQuery(con, sprintf("SELECT [Red (I)], [Green (I)], [Blue (I)] FROM [%s]", "Image"))
georef <- 
  sqlQuery(con, sprintf("SELECT TOP 1 PixelsByX([%s]) AS [ncol], PixelsByY([%s]) AS [nrow] FROM [%s]", "Image", "Image", "Image"))

r0 <- raster(xmn = 0, ymn = 0, xmx = georef$ncol, ymx = georef$nrow, nrow = georef$nrow, ncol = georef$ncol, crs = "+proj=laea")
r <- setValues(brick(r0, r0, r0), as.matrix(colors))
plotRGB(r, asp = 1, interpolate = TRUE)
``` 


If I ask what tables are available, it's not comprehensive. 

```{r}
sqlTables(con)
```

I can read from `LoneTable`, `DrawingOwnedTable`, and `QuerySurface` unsurprisingly. 

```{r}
sqlQuery(con, "SELECT * FROM [LoneTable]")
sqlQuery(con, "SELECT [ID], [X (I)], [Latitude (I)] FROM [DrawingOwnedTable]")
sqlQuery(con, "SELECT [Height (I)], [Center Y (I)] FROM [QuerySurface]")  ## SELECT TOP 1 * [From QuerySurface]

```

But I can also read from `Drawing` and `Surface`. 

```{r}
sqlQuery(con, "SELECT [ID] FROM [Drawing]")
sqlQuery(con, "SELECT TOP 2 [Easting (I)] FROM [Surface]")
```

This is fine except that I can't programmatically get access to the
existence of `Surface` via ODBC, whereas with the Drawing, Query and
stand alone Table I only need list them from `sqlTables` and query the
column names in turn to get a summary of the contents in the .map
file.

I also cannot tell the type of any of the tables without prior knowledge, is that right? I can try for Geom (I) and Surface virtual table columns, but there's no guarantee that any result 
Is this a limitation of the ODBC interface, or one of the RODBC
package, or one of Manifold's ODBC driver?



Thanks!


```{r}
odbcClose(con)
```


# Notes

```R
#' ODBC connection to Manifold map files.
#' 
#' Create an ODBC connection for Manifold GIS. 
#' 
#' See \code{\link[RODBC]{odbcDriverConnect}}
#' @param mapfile character string, path to Manifold project *.map file
#' @param unicode logical
#' @param ansi logical
#' @param opengis logical
#' @details See the documentation for the underlying driver:
#' \url{http://www.georeference.org/doc/using_the_manifold_odbc_driver.htm}
#' @examples
#' \dontrun{
#' f <- system.file("extdata", "AreaDrawing.map", package = "manifoldr")
#' con <- odbcConnectManifold(f)
#' tab <- RODBC::sqlQuery(con, "SELECT * FROM [Drawing]")
#' ## drop [Geom (I)] and give a summary
#' summary(subset(tab, select = -`Geom (I)`))
#' 
#' ## issue a spatial query
#' qtx <- "SELECT [ID], [Name], [Length (I)] AS [Perim], 
#'      BranchCount([ID]) AS [nbranches] FROM [Drawing Table]"
#' sq <- RODBC::sqlQuery(con, qtx)
#' sq
#' }
#' @return RODBC object
#' @importFrom RODBC odbcDriverConnect
#' @importFrom tools toTitleCase 
#' @export
odbcConnectManifold <- function (mapfile, unicode = TRUE, ansi = TRUE, opengis = TRUE)
  
{
  
  full.path <- function(filename) {
    
    fn <- chartr("\\", "/", filename)
    
    is.abs <- length(grep("^[A-Za-z]:|/", fn)) > 0
    
    chartr("/", "\\", if (!is.abs)
      
      file.path(getwd(), filename)
      
      else filename)
    
  }
  unicode <- tools::toTitleCase(tolower(format(unicode)))
  ansi <- tools::toTitleCase(tolower(format(ansi)))
  opengis <- tools::toTitleCase(tolower(format(opengis)))
  
  parms <- sprintf(";Unicode=%s;Ansi=%s;OpenGIS=%s;DSN=Default", unicode, ansi, opengis)
  
  con <- if (missing(mapfile))
    
    "Driver={Manifold Project Driver (*.map)};Dbq="
  
  else {
    
    fp <- full.path(mapfile)
    
    paste("Driver={Manifold Project Driver (*.map)};DBQ=",
          
          fp, ";DefaultDir=", dirname(fp), parms, ";", sep = "")
    
  }
RODBC::odbcDriverConnect(con)
  
}

```
