---
title: "Discovering tables in Manifold via ODBC"
author: "Michael Sumner"
date: "20 January 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(manifoldr)
library(RODBC)
library(raster)
```


In this Project "Components.Map" I have a number of components, to try
different kinds of queries. I am using R with the CRAN package `RODBC`
and my in-development package `manifoldr`.

```{r}
library(RODBC)  ## install.packages("RODBC")
library(manifoldr) ## https://github.com/mdsumner/manifoldr
mapfile <- "C:/temp/Components.map"
con <- odbcConnectManifold(mapfile)
print(con)  ## summary of the connection
```

```{r, echo = FALSE, eval = TRUE}
library(manifoldr)
library(RODBC)
library(raster)
colors <- sqlQuery(con, sprintf("SELECT [Red (I)], [Green (I)], [Blue (I)] FROM [%s]", "Image"))
georef <- 
  sqlQuery(con, sprintf("SELECT TOP 1 PixelsByX([%s]) AS [ncol], PixelsByY([%s]) AS [nrow] FROM [%s]", "Image", "Image", "Image"))

r0 <- raster(xmn = 0, ymn = 0, xmx = georef$ncol, ymx = georef$nrow, nrow = georef$nrow, ncol = georef$ncol, crs = "+proj=laea")
r <- setValues(brick(r0, r0, r0), as.matrix(colors))
plotRGB(r, asp = 1, interpolate = TRUE)
``` 


If I ask what tables are available, it's not comprehensive. 

```{r}
sqlTables(con)
```

I can read from `LoneTable`, `DrawingOwnedTable`, and `QuerySurface` unsurprisingly. 

```{r}
sqlQuery(con, "SELECT * FROM [LoneTable]")
sqlQuery(con, "SELECT [ID], [X (I)], [Latitude (I)] FROM [DrawingOwnedTable]")
sqlQuery(con, "SELECT [Height (I)], [Center Y (I)] FROM [QuerySurface]")  ## SELECT TOP 1 * [From QuerySurface]

```

But I can also read from `Drawing` and `Surface`. 

```{r}
sqlQuery(con, "SELECT [ID] FROM [Drawing]")
sqlQuery(con, "SELECT TOP 2 [Easting (I)] FROM [Surface]")
```

This is fine except that I can't programmatically get access to the
existence of `Surface` via ODBC, whereas with the Drawing, Query and
stand alone Table I only need list them from `sqlTables` and query the
column names in turn to get a summary of the contents in the .map
file.

I also cannot tell the type of any of the tables without prior knowledge, is that right? I can try for Geom (I) and Surface virtual table columns, but there's no guarantee that any result 
Is this a limitation of the ODBC interface, or one of the RODBC
package, or one of Manifold's ODBC driver?



Thanks!


```{r}
odbcClose(con)
```
